Human computational hysteresis: measuring inter-task effects

### 2014-09-23 docs/publications/2014.09_SCIENCE/man.tex
backup work on manuscript

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible to submit batches of small tasks to a large pool of human workers, 
who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Originally used to distribute clerical work, these platforms 
increasingly serve as a fast and cheap means to engage experimental 
participants in a research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
Typical tasks include flagging inapropriate images, annotiting sentiment in 
text excerpts, or judging the relevancy of search results.

The task requester can interact with the platform like a compute server, 
seamlessly 
integrating human and machine computation.  Researchers have put forward the 
term HPU (Human co-Processing Unit), viewing the introduction of 
microtask platforms as a new computing architecture
\cite{5543192}.

Here, we highlight an important way in which HPUs differ from CPUs, with 
serious implications for the design of tasks.  It is well known that people 
are subject to priming effects 
\cite{BJOP1796,No2007,beller1971priming} and, in particular, task-repetition effects
\cite{Gass1999549,sohn2001task}.  
We investigate the effect that previously completed tasks have on workers'
responses during subsequent ones. We call such effects 
\textit{inter-task priming}.  Inter-task priming would amount to a kind of
\textit{hysteresis}, meaning that HPU output is not only a function of the 
current input, but also of the history of inputs.

There has been considerable investigation into the factors that affect the 
quality and nature of microtask work.  These include the level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Research has also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated the 
effects of inter-task effects in microtask work.

To characterize inter-task effects, we present a novel, generalized definition 
of \textit{HPU priming}. Although priming is traditionally used to elucidate 
internal 
psychological mechanisms, our definition characterizes a computational 
property of the HPU, and corresponds to the worst-case bias introduced into an 
algorithm ``running'' on primed HPUs. This definition enables a direct, 
quantititive comparison of HPU priming effects accross domains and 
applications, while offering a practical, intuitive interpretation. We 
demonstrate that machine learning tools can be used to detect priming, an 
approach that could be used to detect priming effects without knowing in 
advance what the nature of the effects might be, nor how priming was imposed.

In the present work, we investigate inter-task priming on the Amazon 
Mechanical Turk platform, using an image-labeling task, one of the most common 
kinds of tasks on that platform 
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}.  
The task requires workers to label images featuring food, table-settings, and 
related utensils and cookware. We divide the set of images shown to a worker 
into an \textit{initial} and \textit{test} set, but, crucially, there was no 
distinction made between these sets, nor any interruption in the flow of tasks 
from the view of the worker. We varied the initial set of images, while 
keeping the test set the same, to analyze the effect that that the initial 
images have on the labels that workers attribute to the test set.

As a point of comparison, we conduct a similar experiment where we induce 
\textit{framing}. Before the workers label images, we indicate that the work 
was funded by a research lab having a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. In both experiments, we 
measure effects using our novel approach, as well as looking for changes in 
the content of labels and label specificity based on an ontology we developed 
for that purpose.

Surprisingly, we found that inter-task effects are much stronger than framing. 
Our results show that initial tasks can significantly alter the focus of 
worker's labels. We also found that inter-task effects can induce workers to 
provide labels of greater specificity. This suggests that careful consideration
should be given to the bundling of tasks when designing a study using a
microtask platform.

Our analysis also highlights important advantages of our novel definition of 
priming.  Our approach incorporates no prior knowledge about the nature
of the task whatsoever.  In contrast, we also sought to detect priming
based on its direct effects on label content and specificity,
drawing on prior knowledge about the images and label semantics.
Remarkably, while being more general, our definition yields a more sensitive
measure of priming when combined with machine learning algorithms. 
This approach could be applied to detect priming when one does not 
\emph{a priori} know whether any priming has occurred, 
nor what the nature of the effects might be.

@@

\paragraph{Experimental setup.}
We solicited 2300 MTurk workers to perform an image-labeling task using images 
that depicted meals, place-settings, and related items such as dishes, and 
kitchenware, and other artifacts naturally arising in dining contexts. Workers 
were randomly assigned into one of four treatments, described below. In all 
treatments, workers were given brief instruc- tions before labeling any images.
In some treatments, workers were then also shown a \textit{framing slide}, 
indicating 
the tasks were funded by a group having a particular research focus. The tasks 
consisted of single images, and workers were asked to provide five descriptive 
labels for each. The last five images that workers labeled, the 
\textit{test set}, were the same for all treatments.

>>>>

\section*{Introduction}
Microtask crowdsourcing platforms like Amazon Mechanical Turk (MTurk) make it 
possible for requesters to submit batches of small tasks to a large pool of 
workers, who do the tasks for fun, a sense of purpose, and remuneration 
\cite{kazai2013analysis,Antin20122925}.  
Originally used to distribute clerical work, these platforms 
increasingly serve as a means to engage experimental 
participants in a research 
setting \cite{paolacci2010running,Berinsky2012351,snow2008cheap,alonso2009can}.
Typical tasks include tagging and categorizing images 
\cite{6116320,Zhai2012357}, transcribing voice recordings 
\cite{chandler2013breaking,paolacci2010running}
or handwritten notes \cite{Berinsky2012351,Finnerty2013}, and judging the 
relevancy of search results \cite{le2010ensuring,grady2010crowdsourcing,alonso2009can,kazai2013analysis}

The task requester can interact with the platform like a compute server, 
seamlessly 
integrating human and machine computation.  Researchers coined the 
term HPU (Human co-Processing Unit), viewing the introduction 
of microtask platforms as a new computing architecture\cite{5543192}.  
Attempts have been made to formalize the HPU instruction set, and libraries 
that provide high-level APIs (Application Programming Interfaces) for writing 
algorithms involving human computation have been developed \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Here, we highlight an important way in which HPUs differ from CPUs, with 
serious implications for any crowdsourcing initiative.  Common sense would 
dictate that people's output during a task will be stochastic: when the same
task is given to different people, or given to the same person at different 
times,
the outputs may, in general, differ.  Stochasticity is not itself problematic,
since the analysis of probabilistic algorithms is fairly 
well-understood. But other non-random effects may also bear on HPU output. 
It is known that people are susceptible to priming effects 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, task-repetition 
effects \cite{Gass1999549,sohn2001task}.

Such repetition effects, or
\textit{inter-task effects}, are neither random, nor are they formally part 
of the instructions given to the HPU.  The ``leakage'' of information from 
recently performed tasks into HPU outputs is a potentially strong source of 
systematic bias, which complicates the the algorithmic view of HPUs. 
Inter-task effects would amount to a kind of \textit{hysteresis}, meaning 
that HPU output is not only a function of the current input, but also of the 
history of inputs.

Here we make two contributions towards understanding inter-task effects
and HPU hysteresis in general.  We present a framework for treating this 
source of bias in the context of HPU algorithms, and we measure the impact of
inter-task effects experimentally in a canonical HPU task: the labeling of 
images.

Using the  Amazon Mechanical Turk (Mturk) platform, we solicited workers to 
perform a series of image-labeling tasks.  
In these tasks, workers labeled images featuring iconic cultural scenes,
food, table-settings, and various utensils and cookware. We divided the set of 
images shown to a given worker into an \textit{initial} and \textit{test} set, 
but, crucially, there was no apparent distinction between these sets from the
perspective of the worker.  We varied the initial set of images, while 
keeping the test set the same, to analyze the effect that that the initial 
images had on the labels that workers attributed to the test set.

There has been considerable investigation into the factors that affect the 
quality and nature of microtask work.  These include such factors as the 
level of 
pay \cite{kazai2013analysis}, training \cite{le2010ensuring}, pre-screening of 
workers \cite{paolacci2010running}, and user-interface design 
\cite{Finnerty2013}.  Researchers have also investigated \textit{framing}, 
by testing the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or using an alternative problem discription
\cite{thibodeau2013natural}.  To our knowledge, no study has investigated 
inter-task effects on microtask platforms.

As a point of comparison, we included treatments in which we induce 
\textit{framing}.
Before the workers label images, we indicate that the work 
was funded by a research lab having a particular area of interest, to see the 
influence this has on the workers’ subsequent labels. 
Surprisingly, inter-task effects were stronger than framing, in some cases 
about twice as strong. 
Our results show that initial tasks can significantly alter the subject 
matter on which worker's labels focus. We also found that inter-task 
effects can induce workers to provide labels of greater or lesser specificity. 
This suggests that careful consideration should be given to the bundling of 
tasks when designing a study using a microtask platform.

To characterize inter-task effects, we present a novel, generalized definition 
of \textit{HPU priming}. Priming is traditionally used to elucidate the
internal psychological mechanisms of cognition.  But here we are concerned with
understanding the outward effects that priming phenomena have on HPU 
algorithms. Thus, before we discuss inter-task effects experimentally, we 
discuss formal definitions for priming as computational phenomenon.  Our
discussion will lead to a natural definition of priming that enables 
comparisons of the severity priming effects accross applications, and enables
the measurement of priming using machine learning tools.

@@

\paragraph{Experimental setup.}

image labeling is common\ldots
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}.

We solicited 2300 MTurk workers to perform an image-labeling task using images 
that depicted meals, place-settings, and related items such as dishes, and 
kitchenware, and other artifacts naturally arising in dining contexts. Workers 
were randomly assigned into one of four treatments, described below. In all 
treatments, workers were given brief instruc- tions before labeling any images.
In some treatments, workers were then also shown a \textit{framing slide}, 
indicating 
the tasks were funded by a group having a particular research focus. The tasks 
consisted of single images, and workers were asked to provide five descriptive 
labels for each. The last five images that workers labeled, the 
\textit{test set}, were the same for all treatments.
